---
title: "Exercise #4"
author: "Leeza M Santiago Millán"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 1
---

# **Set-up libraries and data**

## Load libraries
```{r load_libraries, message=FALSE}
library(tidyverse)
library(pheatmap)
```

## Load data
```{r load_data}
load("data/Rmarkdown_data.Rdata")
```

# **Top 20 significant genes**

## Get names of top 20 significant genes
```{r gene_names}
top20_sigOE_genes <- res_tableOE_tb %>% 
  arrange(padj) %>% 	#Arrange rows by padj values
  pull(gene) %>% 		#Extract character vector of ordered genes
  head(n=20)	
```

## Normalize counts for top 20 significant genes
```{r normalize_counts}
top20_sigOE_norm <- normalized_counts %>%
  filter(gene %in% top20_sigOE_genes)
```

## Gather the columns to have normalized counts to a single column
```{r gather_as_column}
gathered_top20_sigOE <- top20_sigOE_norm %>%
  gather(colnames(top20_sigOE_norm)[2:9], key = "samplename", value = "normalized_counts")
gathered_top20_sigOE <- inner_join(mov10_meta, gathered_top20_sigOE)
```

## Plot using ggplot2
```{r plot}
ggplot(gathered_top20_sigOE) +
  geom_point(aes(x = gene, y = normalized_counts, color = sampletype)) +
  scale_y_log10() +
  xlab("Genes") +
  ylab("log10 Normalized Counts") +
  ggtitle("Top 20 Significant DE Genes") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  theme(plot.title = element_text(hjust = 0.5))
```


Figure 1. Normalized counts for top twenty significant genes in control (orange), knockdown (green), and overexppressed (blue) backgrounds.

# **Create a heatmap of the differentially expressed genes**

## Extract normalized expression for significant genes from the OE and control samples (2:4 and 7:9)
```{r extract_significantgenes}
res_tableOE_tb_sig <- res_tableOE_tb %>%
  filter(padj < 0.05)
```

## Return the normalized counts for the significant DE genes
```{r normalizedcount_DEgenes}
norm_OEsig <- normalized_counts %>% 
  filter(gene %in% res_tableOE_tb_sig$gene) 

meta <- mov10_meta %>%
column_to_rownames("samplename") %>%
data.frame()
```

## Run pheatmap using the metadata data frame for the annotation
```{r heatmap}
pheatmap(norm_OEsig[2:9], 
         cluster_rows = T, 
         show_rownames = F,
         annotation = meta, 
         border_color = NA, 
         fontsize = 10, 
         scale = "row", 
         fontsize_row = 10, 
         height = 20)
```

Figure 2. Comparison of normalized expression of the top twenty significant genes in control, knockdown, and overexppressed backgrounds by heatmap.


# **Session info**
```{r session_info}
sessionInfo()
```
